\chapter{Applications of Cryptographic Hash Functions}
\label{chap:hash_applications}

This chapter is devoted to applications of \glsfirstplural{hash function}.



\section{Ethereum Addresses}

This section assumes some knowledge of the \gls{ethereum} blockchain and
\glspl{elliptic curve}.

\subsection{Address Definition}

We recall that \gls{ethereum} uses \textsc{secp256k1}
for its \gls{ecc};
its private keys are 256-bit nonnegative (unsigned) integers
(32 bytes in length; 64 hexadecimal characters).
Each private key $k$ produces a public key

\begin{equation}
    P = k\cdot G;
\end{equation}

\noindent
here, $G$ is the specified base point of the \gls{elliptic curve}
(the generator for the \gls{cyclic group}).
Uncompressed, the public key is two 256-bit unsigned integers
(64 bytes).
With the compressed or uncompressed designation byte,
a public key is either 65 bytes uncompressed or 33 bytes compressed.

This is a significant amount of space.
In order to save space, \gls{ethereum}~\cite[Eq.~314]{EthereumYellowpaper}
defines the address of every public key.
This is computed by performing the \Keccak{} hash of the public key
and using the rightmost 20 bytes.

\subsection{Colliding Addresses}

While each private key corresponds to a unique public key,
using only 20 bytes of a hash output reduces the unique
number of addresses to $2^{160}$.
Because there are approximately $2^{256}$ unique private and public keys,
this means there are \emph{many} public keys with the same address;
a rough estimate would be that there are approximately $2^{96}$ public keys
per address.

The good news is that, due to \Keccak{} being a strong \gls{hash function}
with collision resistance,
finding any 2 public keys which map to the same address would require
the computation of approximately $2^{80}$ hash values.
Although large, it \emph{may} be possible for this many hash operations
to be computed by a \emph{well-funded and dedicated} adversary.

In practice, an adversary would want to find a private key
corresponding to a \emph{particular} address.
This converts a collision attack into a second-preimage attack.
This instead requires approximately $2^{160}$ hash operations
to find a valid public key.
Considering the belief that performing $2^{128}$ operations is impractical,
performing $2^{160}$ operations is also impractical.
Thus, there is little concern for colliding addresses in \gls{ethereum}
even though they certainly exist.

\subsection{Example}

\begin{example}[Ethereum Address]
\exampleCodeReference{examples/hash\_applications/ethereum\_addr.py}

We have the private key

\begin{align}
    k &= 2^{248} + 1 \nonumber\\
&= \texttt{0100000000000000000000000000000000000000000000000000000000000001};
\end{align}

\noindent
the second line lists the private key in hexadecimal form.
The corresponding public key in hexadecimal form is

\begin{align}
    &P = k\cdot G = \nonumber\\
    &(\texttt{e4dbb4350d84eabec1d67e40a398a78a8e6d719d86914393fca83b88dbe927af},
        \nonumber\\
    &\texttt{b80fe66bf659859889a544623c945d0bd80d855f649e8c197be3aa41fe0390f8}).
\end{align}

\noindent
By computing the \Keccak{}-256 hash of this and taking the rightmost
20 bytes, we see

\begin{equation}
    \texttt{addr} = \texttt{7691ee0343b9a529675e1a8a70197b3b704f90b7}.
\end{equation}

\noindent
This address is much more manageable than the public key.
\end{example}



\section{Commitment Schemes}

A \emph{commitment scheme} is a method to commit to a specific value
without revealing any information about the value;
at a later point in time, it is possible to reveal the committed
value and prove that this particular value was the commitment.
Ideally, a commitment scheme is \emph{hiding} and \emph{binding}:
\emph{hiding} because no information about the committed value is revealed;
\emph{binding} because the committed value cannot be changed after
it is committed.

Suppose we have our \gls{hash function} $H:\braces{0,1}^{*}\to\braces{0,1}^{n}$.
If $m\in\braces{0,1}^{*}$ is the value to be committed,
we choose our random value $r\chooseRandom{}\braces{0,1}^{n}$.
The commitment is then

\begin{equation}
    c \mathDef{} H(r\|m).
\end{equation}

\noindent
At a future point in time, revealing the value $m$ and the random value $r$
will show that $m$ was the value previously committed.
This is due to the fact that it is difficult to determine
$r'\in\braces{0,1}^{n}$ and $m'\in\braces{0,1}^{*}$
such that $c = H(r'\|m')$.

\begin{example}[Commitment Scheme]
\exampleCodeReference{examples/hash\_applications/commitment\_scheme.py}

We look at a commitment scheme using \MDFive{};
see Listing~\ref{list:commitment_scheme_md5}.
We are seeking to commit to the byte value

\begin{equation}
    m \mathDef{} \texttt{01}.
\end{equation}

\noindent
The random value $r$ is the \MDFive{} hash of 16 zero bytes:

\begin{equation}
    r \mathDef{} \texttt{4ae71336e44bf9bf79d2752e234818a5}.
\end{equation}

\noindent
The resulting commitment $c$ is then

\begin{align}
    c &\mathDef{} \text{\MDFive{}}\parens{r\|m} \nonumber\\
        &= \texttt{ff95213ae708e6443c5f0ae2d8f5fe41}.
\end{align}

\noindent
At a later point in time, $m$ and $r$ may be revealed.

\input{code/hash_applications/commitment_scheme_md5.tex}

\emph{Note well:} We note that in an actual commitment scheme,
such a random value \emph{should not be used};
the value should be truly random or drawn from a \glsfirst{csprng}.
\end{example}

It is not necessary that $r$ be a bit string of length $n$
in the commitment scheme;
this was chosen for convenience.
Even so, it must be long enough so that it is impractical
to test every possible value for $r$ and $m$.
If the random value $r$ was not included in the commitment,
problems may arise.
If the commitment was just $H(m)$,
it may be possible to compute all possible values
and determine which value is committed;
thus, the scheme would not be hiding.



\section{Hash Chains}

Hash chains use \glspl{hash function} recursively.
If $H$ is a \gls{hash function}, then a hash chain of $x$ is

\begin{equation}
    x, H(x), H(H(x)), H(H(H(x))), H(H(H(H(x)))), \cdots
\end{equation}

\noindent
This can be more compactly written as

\begin{equation}
    H^{0}(x), H^{1}(x), H^{2}(x), H^{3}(x), H^{4}(x), \cdots
\end{equation}

\noindent
where we have

\begin{equation}
    H^{n}(x) = \begin{cases}
            x\quad &n = 0 \\
            H(H^{n-1}(x))\quad &n \ge 1 \\
    \end{cases}.
\end{equation}

\noindent
Depending on the circumstance, the initial value ($x$ or $H^{0}(x)$)
may not be included.

Hash chains enable easy iteration in the forward direction.
By using a secure \gls{hash function}, it is impractical to iterate in reverse,
as doing so would require computing \gls{hash function} preimages.

\begin{example}[Hash Chain]
\exampleCodeReference{examples/hash\_applications/hash\_chain.py}

We look at a hash chain using \MDFive{};
see Listing~\ref{list:hash_chain_md5}.
The initial byte array is

\begin{align}
    x &= H^{0}(x) \nonumber\\
        &= \texttt{00112233445566778899aabbccddeeff}.
\end{align}

\input{code/hash_applications/hash_chain_md5.tex}

\noindent
Iterating \MDFive{} five times gives us the value

\begin{equation}
    H^{5}(x) = \texttt{e94712ae81bf26d81b0e7475fbf6b3f8}.
\end{equation}
\end{example}


% Sections on Merkle Trees
\input{tex/hash_app_merkle.tex}


\section{Hash-based Message Authentication Code}
\label{sec:hash_hmac}

As mentioned in Chapter~\ref{sec:symmetric_mac},
\glspl{mac} are used for message authentication;
Bob can be certain Alice sent him the message and that it was not
corrupted in transit.
One method for implementing a \gls{mac} is based on a \gls{hash function},
hence \gls{hmac} (HMAC)~\cite{HMAC1996,rfc2104}.
Any \gls{hash function} may be used,
but \glspl{hmac} based on \MDFive{} or \ShaOne{} should \emph{only}
be used if \emph{required}; see~\cite{rfc6151}.

Further discussion about \gls{hmac} is relegated to
Appendix~\ref{app:crypto_hmac}.



\section{HMAC-based Key Derivation Function}
\label{sec:hash_hkdf}

We mentioned in Chapter~\ref{sec:kdf} that \glspl{kdf}
take initial key material which
has sufficient entropy (but which may not be uniform)
and converts it into a uniform cryptographically-strong key.
One way this can be done is through the use of a \gls{hash function}
via the HMAC construction; see~\cite{HKDF2010,rfc5869}.
In this case, it is called an \gls{hkdf} (HKDF).
The main idea is to first compress the available entropy;
this compressed entropy is then expanded into cryptographic keys.
For instance, it could be used to derive an encryption key
from a \gls{shared secret} between Alice and Bob.

Further discussion about \gls{hkdf} is relegated to
Appendix~\ref{app:crypto_hkdf}.



\section{Mask Generation Functions and Extendable Output Functions}
\label{sec:hash_mgf_xof}

A \gls{mgf} is like a \gls{hash function}
except it has a variable digest length;
in particular, the length may be made arbitrarily long.
There are times when this is useful, especially if the digest
of a \gls{hash function} is too short.
A similar concept is an \gls{xof}.

Further discussion about \glspl{mgf} and \glspl{xof} is relegated to
Appendix~\ref{app:crypto_additional_uniform_bytes}.



\section{Password Hashing}
\label{sec:hash_apps_password_hashing}

We briefly describe another application of \glsfirstplural{hash function}:
safe password storage.
As mentioned in Chapter~\ref{sec:kdf}, \glspl{kdf}
are also used for protecting passwords.

Password storage is difficult.
One way to store passwords is to not store the password \emph{itself}
but to rather store its \emph{hash}.
In order to make each password hash unique, each password
is also given a random \emph{\gls{salt}}.
A \gls{salt} is a bit string which is uniformly random used
for domain separation;
the proper length will be discussed below.

If a \gls{salt} is not used (or a \gls{salt} is repeated),
then it is possible to make \emph{rainbow tables}:
find a list of the most common passwords and hash them,
either by themselves or together with the repeated \gls{salt}.
This is why it is important to not reuse \glspl{salt}:
reusing \glspl{salt} makes it more efficient to crack multiple passwords.
Using a repeated \gls{salt} is almost as bad as
using no \gls{salt} at all.
This is also why it is important to use \glspl{salt}
of sufficient length and entropy:
it should be impractical to make a rainbow table
of every possible \gls{salt} value,
and it should not be possible to \emph{derive} the \gls{salt}
from related information.

While \glsfirstplural{hash function} like \ShaTwo{} or \ShaThree{} may be used,
when security is important, it is important to use a \gls{hash function}
\emph{specifically designed} for password hashing.
Some examples are \texttt{scrypt}~\cite{scryptPaper,rfc7914},
\texttt{Argon2}~\cite{argon2},
and balloon hashing~\cite{cryptoeprint:2016:027}.
When using these \glspl{hash function}, it is possible to tune
the difficulty parameter so that users are only mildly affected
by password validation,
but adversaries have a much more difficult time
when attempting to brute-force potential passwords.
The difficulty parameter ensures that password cracking is difficult
even when performed offline.
We note that even though PBKDF2~\cite{rfc8018} is recommended
as of 2017, it was designed in 2000~\cite{rfc2898}
and should not be used due to the insufficient protection
it supplies~\cite{blocki2018economics}.
In~\cite{blocki2018economics}, the authors recommend \emph{not}
using \texttt{bcrypt}~\cite{bcryptPaper}
and PBKDF2~\cite{rfc2898}.
The \texttt{bcrypt} \gls{kdf}
is derived from the Blowfish cipher~\cite{BlowfishAlg}.
Both PBKDF2 and \texttt{bcrypt} allow for varying the computation time,
but the fact that they do not have large memory requirements
enables the construction of dedicated hardware
for inexpensive attacks.

There have been a number of instances where user passwords have been
hacked; see the examples in~\cite{blocki2018economics}
and Figure~\ref{fig:xkcd_encryptic}.
Suggestions for strong passwords may be found in
Figure~\ref{fig:xkcd_password_strength}.
As we can see from Figure~\ref{fig:xkcd_password_reuse},
password reuse is insecure.

\input{figures/xkcd/xkcd_encryptic.tex}
\input{figures/xkcd/xkcd_password_strength.tex}
\input{figures/xkcd/xkcd_password_reuse.tex}

\subsection*{\Glsentrytext{salt} Length}

It is important that a \gls{salt} be long enough so that it is
impractical to form rainbow tables.
Because of this, \glspl{salt} should be at least be 128 bits (16 bytes).
Ideally, \glspl{salt} should be 256 bits (32 bytes).
Salts may be longer, but the \emph{purpose} of a \gls{salt}
is to ensure a unique output even if the same password is used multiple times.
To make the password derivation more difficult,
use a \gls{hash function} designed for password hashing and
choose the appropriate difficulty setting with iteration count
and memory requirements.



\section{Concluding Discussion of Hash Functions}

There has been a lot of information presented here on \glspl{hash function}.
From all the applications discussed, it is obvious
that \glspl{hash function} are used all over cryptography.
Also, see the discussion of geohashing in Figure~\ref{fig:xkcd_geohashing}.

\input{figures/xkcd/xkcd_geohashing.tex}
