\section{Groups}
\label{sec:math_groups}

\subsection{Why do we care about Groups?}
\Glspl{group} arise in many different areas.
In cryptography, we will encounter them in the \gls{dhke}
and \glspl{signature}.
For \glspl{group} to be used in practice, we will need to encode them;
that is, we will need to know how to represent them on a computer.

\subsection{Intuition and Examples}
\emph{\Glspl{group}} are mathematical objects which we will
frequently encounter.
Informally, a \gls{group} is a generalization of the integers under addition,
so we start by looking at them first.

\begin{example}[Integers under addition]
The standard example of a \gls{group} is to consider the
integers $\Z$ under addition.

Given $a,b\in\Z$, we know that $a+b\in\Z$.
This means that the integers are \emph{closed} under addition:
adding two integers will always give us another integer.

Addition is also \emph{\gls{associative}}.
That is, for $a,b,c\in\Z$, we have

\begin{equation}
    \parens{a + b} + c = a + \parens{b + c}.
\end{equation}

\noindent
In this way, we do not have to worry about which way we add integers;
we will always get the same result regardless of order.

There is also a special integer $0\in\Z$.
It is special in that adding $0$ to $a$ gives us $a$:

\begin{equation}
    0 + a = a + 0 = a.
\end{equation}

\noindent
No other integer has this property.

We also know the (additive) inverse of $a$: $-a$.
This is because

\begin{equation}
    a + (-a) = 0.
\end{equation}

\noindent
It also turns out that the (additive) inverse is \emph{unique}.

We note that every integer can be written as
repeated additions of $1$.
If $a\ge0$, then we have

\begin{equation}
    a = \underbrace{1 + 1 + \cdots + 1}_{\text{$a$ times}} \quad a\ge 0.
    \label{eq:math_groups_integers_1_gen_pos}
\end{equation}

\noindent
For this to apply to negative numbers, we allow for adding $1$'s
additive inverse $-1$.
If $a<0$, then

\begin{equation}
    a = \underbrace{(-1) + (-1) + \cdots + (-1)}_{\text{$\abs{a}$ times}}
        \quad a < 0.
    \label{eq:math_groups_integers_1_gen_neg}
\end{equation}

\noindent
In this way, $1$ \emph{generates} the \gls{group} of integers under addition.

We also notice that addition is \emph{\gls{commutative}}:
given $a,b\in\Z$, we have

\begin{equation}
    a + b = b + a.
\end{equation}

\noindent
Although not all \glspl{group} have this property, the ones we care about do.
\end{example}

\begin{example}[Rationals under addition]
Similar to the integers under addition, the rational numbers $\Q$
under addition also form a \gls{group}.

Given $a,b\in\Q$ with $a = p/q$ and $b = r/s$,
we know

\begin{align}
    a + b &= \frac{p}{q} + \frac{r}{s} \nonumber\\
        &= \frac{ps + qr}{qs}\in\Q.
\end{align}

\noindent
This shows us that the rationals are closed under addition.

We know that addition of rationals is \gls{associative} and \gls{commutative}.
Furthermore, $0\in\Q$ is the identity element.
If $a = \frac{p}{q}$ as above, then the inverse is

\begin{equation}
    -a = \frac{-p}{q}
\end{equation}

\noindent
because

\begin{align}
    a + \parens{-a} &= \frac{p}{q} + \frac{-p}{q} \nonumber\\
        &= \frac{pq - pq}{q^{2}} \nonumber\\
        &= 0.
\end{align}

\noindent
This shows that the rationals under addition form a \gls{group}.
\end{example}

\begin{example}[Integers under modular addition]
We now look at another standard example: integers under modular arithmetic.
Let $n > 1$.
Then

\begin{equation}
    \Z_{n} = \braces{0, 1, \cdots, n-1}.
\end{equation}

\noindent
For $a,b\in\Z_{n}$, we know there is some $c\in\Z_{n}$ such that

\begin{equation}
    a + b = c \mod n.
\end{equation}

\noindent
This shows us that $\Z_{n}$ is closed under addition.

We know $0\in\Z_{n}$ and we always have

\begin{equation}
    0 + a = a + 0 = a
\end{equation}

\noindent
for $a\in\Z_{n}$.
Thus, $0$ is the additive identity.

Given $a\in\Z_{n}\setminus\braces{0}$, we know

\begin{equation}
    a + \parens{n-a} = 0 \mod n,
\end{equation}

\noindent
Thus, the additive inverse of $a$ is $n-a$;
$0$ is its own additive inverse.
Therefore, we see that $\parens{\Z_{n},+}$ is a \gls{group}.
\end{example}

\begin{example}[Positive Rationals under multiplication]
Similar to the integers under addition, the positive rational
numbers $\Q^{+}$ under multiplication form a \gls{group}.
In this case, the multiplicative identity is $1$.
\end{example}

\begin{example}[Nonexample: Natural numbers under addition]
We know that the natural numbers under addition are
very similar to the integers under addition.
We know that we can always add two natural numbers,
so that $a,b\in\N$ implies $a+b\in\N$;
thus, the naturals are closed under addition.
Addition is also \gls{associative} and \gls{commutative}.

We also have the identity $0\in\N$, so

\begin{equation}
    a + 0 = a
\end{equation}

\noindent
for all $a\in\N$.

Unfortunately, we do not have additive inverses;
this is because the natural numbers do not include the negative integers.
That is, for $a\in\N\setminus\braces{0}$, there is no $b\in\N$ such that

\begin{equation}
    a + b = 0.
\end{equation}

\noindent
Thus, $\parens{\N,+}$ is \emph{not} a \gls{group}.
\end{example}

\subsection{Formal Definition}

\begin{defn}[Group]
A \gls{group} is a \gls{set} $G$ together with a binary operation $\cdot$
such that, given $a,b\in G$, $a\cdot b\in G$;
that is, $G$ is closed under the binary operation $\cdot$.

The binary operation also satisfies the following properties:

\begin{itemize}
\item Associativity: for all $a,b,c\in G$, we have

\begin{equation}
    \parens{a\cdot b}\cdot c = a\cdot \parens{b\cdot c}.
\end{equation}

\item Identity element: there is an element $e\in G$ such that
    for all $a\in G$,

\begin{equation}
    a\cdot e = e\cdot a = a.
\end{equation}

\item Inverses: for every $a\in G$ there is a $b\in G$ such that
    
\begin{equation}
    a\cdot b = b\cdot a = e.
\end{equation}

\noindent
We henceforth denote the inverse of $a$ as $a^{-1}$.
\end{itemize}
\end{defn}

Formally, we write that $\parens{G,\cdot}$ is a \gls{group}.
Informally, we write that $G$ is a \gls{group} when the operation is understood.
Additionally, we frequently write $ab$ for $a\cdot b$.

\subsection{Continued Discussion}
Although the generic group operation was listed as ``multiplication'',
we could very well have used addition.
We will sometimes refer to some \glspl{group} as ``multiplicative groups''
while others as ``additive groups'';
this just means we use the multiplication sign $\cdot$
or the addition sign $+$ to denote the group operation.
Nothing changes, as we can use both symbols,
but there are conventions.
In additive groups, the identity element is usually denoted $0$;
in multiplicative groups, the identity element is usually denoted $1$.

Because $\parens{\Z,+}$ is a \gls{group}, we know that \glspl{group} can have
an infinite number of elements.
In what follows, we will be particularly interested
in \emph{\glspl{finite group}}:
\glspl{group} that have a finite number of elements.
We let $\abs{G}$ denote the order of the \gls{group}
(the number of elements in $G$).

At times, we may drop explicit reference to the operation if it is understood.
For instance, we may refer to $\Z$ as a \gls{group},
when more formally we should say $\parens{\Z,+}$ is a \gls{group}
in order to reference both the \emph{\gls{set}} (the integers $\Z$)
and the \emph{operation} (integer addition).
Even so, this is long, tedious, and painful, so we will,
at times, forego the formalities when the operation is clear
from the context.

\subsection{More Definitions}
Let $G$ be a \gls{group} and $H \subseteq G$;
more explicitly, $\parens{G,\cdot}$ is a \gls{group} and $H$ is a subset of $G$.
We say that $H$ is a \emph{\gls{subgroup}} of $G$, written as $H \le G$,
if $H$ with the group operation inherited from $G$ is also a \gls{group}.
Every \gls{group} $G$ has at least two \glspl{subgroup}:
$G$ (the entire group) and $\braces{e}$
(the group that is just the identity element $e$).
A \gls{subgroup} is called a \emph{proper subgroup} when
$H\le G$ and $H\ne G$.
If $H$ is a proper subgroup, then we may write $H < G$.
The \gls{group} $\braces{e}$ is called the \emph{trivial} subgroup.

\begin{example}[Subgroup Example: $2\Z \le \Z$]
We let

\begin{equation}
    2\Z \mathDef{} \braces{\ldots, -4, -2, 0, 2, 4, \ldots};
\end{equation}

\noindent
that is, $2\Z$ is the set of even integers.
Naturally, we have $2\Z\subseteq\Z$.
For every $a\in2\Z$ there is some $b\in\Z$ so that $a = 2b$.

Given $x,y\in2\Z$, we set $x = 2m$ and $y = 2n$ for $m,n\in\Z$.
We then see

\begin{align}
    x + y &= 2m + 2n \nonumber\\
        &= 2\parens{m+n}\in2\Z.
\end{align}

\noindent
Thus, $2\Z$ is closed under addition.
Because addition on $\Z$ is \gls{associative},
addition on $2\Z$ is also \gls{associative}.

Because $0\in\Z$ and

\begin{equation}
    0 + x = x\in2\Z,
\end{equation}

\noindent
we still have the same identity element $0\in2\Z$.

We let $x\in2\Z$ with $x = 2m$ for $m\in\Z$.
If $y = -2m$, then $y\in\Z$ and

\begin{align}
    x + y &= 2m + \parens{-2m} \nonumber\\
        &= 0\in2\Z.
\end{align}

\noindent
Thus, we see that every element in $2\Z$ has an additive inverse
within $2\Z$.
We have just shown that $\parens{2\Z,+}$ is a \gls{group};
because $2\Z\subseteq\Z$, $2\Z$ is a \gls{subgroup} of $\Z$.

In general, for $n\in\N\setminus\braces{0}$ we let

\begin{equation}
    n\Z \mathDef{} \braces{\ldots, -2n, -n, 0, n, 2n, \ldots}.
\end{equation}

\noindent
Then $n\Z\le\Z$.
This is a \emph{proper} subgroup when $n\ge2$.
\end{example}

For a multiplicative group $\parens{G,\cdot}$ and $g\in G$, we define

\begin{equation}
    \angles{g} \mathDef{} \braces{g^{n} \mid n\in\Z}.
\end{equation}

\noindent
In this case, we see that $\angles{g}$ contains all group elements
of the form $g^{n}$ for $n\in\Z$.
This is similar to Eqs.~\eqref{eq:math_groups_integers_1_gen_pos}
and \eqref{eq:math_groups_integers_1_gen_neg}.
Here, we have

\begin{equation}
    g^{n} = \underbrace{g \cdot g \cdots g}_{\text{$n$ times}} \quad n\ge 0
\end{equation}

\noindent
and

\begin{equation}
    g^{n} = \underbrace{g^{-1} \cdot g^{-1} \cdots g^{-1}}_{\text{$\abs{n}$
        times}} \quad n < 0.
\end{equation}

\noindent
In this case, $\angles{g}$ is the \gls{subgroup} of $G$ which consists
of products of $g$ or $g^{-1}$;
this allows us to write $\angles{g}\le G$.
We say $G$ is a \emph{\gls{cyclic group}} if there is a $g\in G$ such that
for all $h\in G$ there is an $n\in\Z$ such that $h = g^{n}$.
In this case, we have $G = \angles{g}$ and 
say that $g$ is a \emph{generator} of $G$.
\Glspl{cyclic group} are important in cryptography because
these \glspl{group} are used in the \gls{dhke};
they are also used to construct \glspl{signature}.

\begin{example}[Cyclic Group: $\Z$]
As we saw previously, $1$ is a generator of the integers $\Z$ under addition.
Thus, we can write $\angles{1} = \Z$.
We note that $-1$ is also a generator, so $\angles{-1} = \Z$.
\end{example}

\begin{example}[Non-cyclic Group: $\Q$]
We can consider the rational numbers $\Q$ under addition.
For any $a \in \Q\setminus\braces{0}$, we know that
and $\frac{a}{2}\in\Q\setminus\braces{0}$.
Even so, we see that $\frac{a}{2} \not\in \angles{a}$.
Thus, $\Q$ is not a \gls{cyclic group}.
\end{example}

In all of the groups we have looked up to this point,
all of our group operations satisfy $a\cdot b = b\cdot a$;
that is, the group operation is \emph{\gls{commutative}}.
We say that $G$ is an \emph{\gls{abelian group}} if for all $a,b\in G$
we have $a\cdot b = b\cdot a$.
This need not always be the case, but we will only consider
\glspl{abelian group} here because the \glspl{group} which normally arise
in cryptography are abelian.

We note that addition and multiplication are familiar
\gls{commutative} operations;
we always have $a+b = b+a$ and $a\cdot b = b\cdot a$ for real numbers.
We know that subtraction and division are \emph{not} \gls{commutative};
in general, we have $a-b \ne b-a$ and $\frac{a}{b} \ne \frac{b}{a}$.

We will usually denote $a\cdot b$ as $ab$ from now on
when there is no cause for confusion.

\subsection{Encoding Groups}

When working with the group $\Z_{n}$,
we can encode $k\in\Z_{n}$ by its binary representation.
Other groups may have different representations.
