\chapter{Secret Sharing Protocols}
\label{chap:secret_sharing}

We now consider the situation where individuals wish to share
secret information.
Perhaps a collection of individuals wish to split a private key
between multiple parties;
this way, a specified threshold will need to work together
to use the private key.

If the reader is not familiar with \gls{lagrange interpolation},
it would be a good idea for him to review the material
in Chapter~\ref{sec:math_lagrange}, as it will be used extensively
in this chapter.
Furthermore, the reader is assumed to also be familiar with
\glspl{bilinear} and BLS signatures as discussed in Chapter~\ref{chap:pairing}.

\emph{Notation Convention:}
Throughout this chapter we will be looking at
$\parens{t,n}$-threshold schemes.
By this we mean that $t+1$ participants are required to work together
in order to learn the secret
and that $t$ individuals working together gain no additional information.
We point this out because there are conflicting conventions
throughout the literature.

\section{The Need for Secret Sharing Protocols}

Suppose that Alice, Bob, and Charlie start a business.
Being equals, they want a majority to be able to withdraw
money for business expenses at any time.
At the same time, they do not fully trust each other.
They would like to split the private key between the three of them
so that 2 out of 3 participants will be required to open the vault.
Alice, Bob, and Charlie will need to use a form of secret sharing
to share their private vault key.
The challenge is that Alice should know
\emph{nothing} about the private key without assistance from either
Bob or Charlie.

We will start by discussing the original secret sharing
protocol (Shamir's Secret Sharing)~\cite{shamir1979share}
in Chapter~\ref{sec:ss_shamir}.
From here, we will allow for the possibility of malicious
actors in Chapter~\ref{sec:ss_verifiable}.
Finally, we discuss the solution to Alice, Bob, and Charlie's
problem when we give an overview of \gls{distributed key generation} (DKG)
in Chapter~\ref{sec:ss_dkg};
the DKG protocol is discussed in Chapter~\ref{sec:ss_dkg_protocol}.
As a follow up to DKG, we discuss how to make group signatures
in Chapter~\ref{sec:ss_threshold}.

\input{tex/secret_sharing_shamir.tex}

\section{Verifiable Secret Sharing}
\label{sec:ss_verifiable}

Verifiable Secret Sharing protocols seek to combat the shortcomings
discussed in Chapter~\ref{ssec:ss_shamir_discussion}
of Shamir's Secret Sharing protocol.
Feldman's scheme~\cite{feldman1987practical}
is frequently used but by itself is not
secure~\cite{gennaro1999secure,gennaro2002revisiting}.

Due to the fact that VSS protocols are used when attempting to distribute
a key between individuals,
we do not mention VSS protocols any further but rather
proceed to discuss \gls{distributed key generation} protocols.

\input{tex/secret_sharing_dkg.tex}
\input{tex/secret_sharing_threshold.tex}
\input{tex/secret_sharing_dkg_examples.tex}
